---
title: "mduennebacke_017"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2024-01-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Preperation

## Loading packages 

```{r loading_packages, message=FALSE, warning=FALSE}

library(protti)
library(tidyverse)
library(data.table)
library(dplyr)
library(magrittr)
library(stringr)
library(tidyr)
library(ggplot2)
library(gridExtra)


```

## Loading data


```{r loading_data, message=FALSE, warning=FALSE}

setwd("/Volumes/biol_bc_picotti_1/Maximilian_D/017_ß-lactamase_Re")

DIA_raw <- read_protti("20240224_120316_mduennebacke_Report_small_library.csv")
```
## filter data

```{r cleaning_data, message = FALSE, warning = FALSE}


DIA_raw$intensity_log2 <- log2(DIA_raw$fg_quantity)


DIA_raw_norm <- normalise(DIA_raw,
  sample = r_file_name,
  intensity_log2 = intensity_log2,
  method = "median"
)

qc_median_intensities(
  DIA_raw_norm,
  r_file_name,
  pep_grouping_key,
  normalised_intensity_log2,
  plot = TRUE,
  interactive = FALSE
)

DIA_clean <- DIA_raw_norm %>%
  filter(eg_qvalue <= 1e-5) %>%
  #filter(pg_fasta_files != "contaminants", pg_fasta_files != "NaN") %>%
  filter(fg_quantity > 10) %>%
  #mutate(intensity_log2 = log2(fg_quantity)) %>%
  filter(pep_is_proteotypic == T)# %>%
  #mutate(uniprot_id = pg_protein_accessions)

qc_median_intensities(
  DIA_clean,
  r_file_name,
  pep_grouping_key,
  normalised_intensity_log2,
  plot = TRUE,
  interactive = FALSE
)




unis <- unique(DIA_clean$pg_protein_accessions) # make vector for fetch_uniprot
```
## Load data from uniprot and join with DIA dataframe

```{r uniprot, message = FALSE, warning = FALSE}

uniprot <- # download protein information from UniProt
  fetch_uniprot(
    unis,
    columns =  c(
      "protein_name",
      "gene_names",
      "length",
      "sequence",
      "xref_pdb"
    )
  ) #%>%
  # rename(
  #   protein_sequence = sequence,
  #   length_protein = length,
  #   uniprot_id = accession
  # )

DIA_clean_uniprot <- DIA_clean %>%
  left_join(uniprot, by = c("pg_protein_accessions" = "accession")) %>% # rejoin with annotations
  find_peptide(sequence, pep_stripped_sequence) %>%
  assign_peptide_type(aa_before, last_aa, aa_after) %>%
  # sequence_coverage(sequence, pep_stripped_sequence) %>%
  distinct()

DIA_clean_uniprot$condrep <- paste(DIA_clean_uniprot$r_condition, DIA_clean_uniprot$r_replicate, sep = "_")

proteins_identified <- uniprot %>%
  distinct(accession)

# fwrite(proteins_identified, file = "identified_proteins_CaM_LiP.csv", sep = ",")


```

#Data analysis


## Volcano plots precurser lvl


```{r Volcano_plot, message = FALSE, warning = FALSE}




Volcano_input <- DIA_clean_uniprot %>%
  unique() %>%
  assign_missingness(r_file_name,
                     r_condition,
                     fg_id,
                     normalised_intensity_log2,
                     ref_condition = "0_mM",
                     retain_columns = c(pg_protein_accessions, pep_grouping_key, start, end, pep_type, sequence, length))


t_test_result_proDA_pep <- calculate_diff_abundance(data = Volcano_input,
                                r_file_name,
                                r_condition,
                                fg_id,
                                normalised_intensity_log2,
                                missingness,
                                comparison = comparison,
                                ref_condition = "0_mM",
                                method = "moderated_t-test",
                                retain_columns = c(pg_protein_accessions, pep_grouping_key, start, end, pep_type, sequence, length))





condition_vulcano <- c("1_mM_vs_0_mM")
t_test_result_proDA_pep_comp <- t_test_result_proDA_pep[t_test_result_proDA_pep$comparison %in% condition_vulcano, ]


volcano_plot(t_test_result_proDA_pep_comp, fg_id, diff, pval, method = "target", target_column = pg_protein_accessions, target = c("P62593") ,  x_axis_label = "log2(fold change) ", title = "Volcano plot precursor level", y_axis_label = "-log10(p-value)", log2FC_cutoff = 1, significance_cutoff = 0.01, interactive = T)


#volcano_plot(t_test_result_proDA_pep_comp, pep_grouping_key, diff, adj_pval, method = "target", target_column = pg_protein_accessions, target = c("P24941") ,  x_axis_label = "log2(fold change)", title = "Volcano plot precursor level", y_axis_label = "-log10(q-value)", log2FC_cutoff = 1, significance_cutoff = 0.05, interactive = T)



```
## fragment traces

```{r traces, fig.align = "center", message = FALSE, warning = FALSE}

# Function to plot based on input sequence
plot_sequence <- function(df, sequence_value) {
  # Filter the dataframe based on the input sequence
  subset_df <- df[df$pep_grouping_key == sequence_value, ]
  
  # Create the plot
  plot <- ggplot(subset_df, aes(x = condrep, y = normalised_intensity_log2, group = 1)) +
    geom_point() +  # Use points for each value
    geom_line() +   # Add lines connecting the points
    labs(title = paste("Intensity Plot for", sequence_value),
         x = "condrep",
         y = "Normalized Intensity (log2)") +
    theme_minimal()
  
  return(plot)
}


plot_sequence(DIA_clean_uniprot, "_KDIGPNKR_.2")


# 
# # Function to plot all sequences for a specific protein accession
# plot_sequences_for_protein <- function(df, protein_accession) {
#   # Filter the dataframe based on the input protein accession
#   subset_df <- df[df$pg_protein_accessions == protein_accession, ]
#   
#   # Create the plot
#   plot <- ggplot(subset_df, aes(x = condrep, y = normalised_intensity_log2, group = pep_grouping_key, color = pep_grouping_key)) +
#     geom_point() +  # Use points for each value
#     geom_line() +   # Add lines connecting the points
#     labs(title = paste("Intensity Plot for Protein Accession:", protein_accession),
#          x = "condrep",
#          y = "Normalized Intensity (log2)") +
#     theme_minimal() +
#     scale_color_discrete()  # Adjust line and point colors
#   
#   return(plot)
# }
# 
# 
# 
# plot_sequences_for_protein(DIA_clean_uniprot, "P54652")

```



## Woods plot

```{r woods, fig.align = "center", message = FALSE, warning = FALSE}

t_test_result_proDA_pep_comp %>%
  filter(pg_protein_accessions == "P62593") %>%
  mutate(significant = ifelse(pval < 0.01, TRUE, FALSE)) %>%
  drop_na() %>%
  woods_plot(diff,
             start,
             end,
             length,
             protein_id = pg_protein_accessions,
             highlight = significant,
             #colouring = pep_type,
             colouring = adj_pval,
             fold_change_cutoff = 1)

```
## digestion of POI

```{r POI_digest, fig.align = "center", message = FALSE, warning = FALSE}

specific_accession <- "P62593"

# Filter the DataFrame for the specific protein accession
filtered_df <- subset(DIA_clean_uniprot, pg_protein_accessions == specific_accession)

# Create a bar plot of the counts of entries in the pep_digest_type_trypsin_p column
plot <- ggplot(filtered_df, aes(x = pep_digest_type_trypsin_p)) +
  geom_bar() +
  labs(title = "Digestion Type Counts for Specific Protein Accession",
       x = "Digestion Type",
       y = "Count")

# Show the plot
print(plot)


```
##Barcode plot

```{r Barcode, message = FALSE, warning = FALSE}

t_test_result_proDA_pep_comp %>%
  filter(pg_protein_accessions == "P62593") %>%
  barcode_plot(start,
               end,
               length,
               cutoffs = c(diff = 1, pval = 0.01)
               )








```

#QC

#CV

```{r quality_control_cvs, fig.width = 12, fig.align= "center",  message = FALSE, warning = FALSE}

qc_cvs(
  data = DIA_clean_uniprot,
  grouping = pep_grouping_key,
  condition = r_condition,
  intensity = fg_quantity,
  plot = FALSE
)

qc_cvs(
  data = DIA_clean_uniprot,
  grouping = pep_grouping_key,
  condition = r_condition,
  intensity = fg_quantity,
  plot_style = "density",
  plot = TRUE
)

qc_cvs(
  data = DIA_clean_uniprot,
  grouping = pep_grouping_key,
  condition = r_condition,
  intensity = fg_quantity,
  plot_style = "violin",
  plot = TRUE
)

```

### Intensity distribution

Intensity distributions are plotted for the whole dataset.

```{r quality_control_log2_int_distr, fig.align= "center", message = FALSE, warning = FALSE}


qc_intensity_distribution(
  DIA_clean_uniprot,
  condrep,
  pep_grouping_key,
  intensity_log2,
  plot_style = "histogram"
)

```

### Missed cleavages


```{r quality_control_missed_cleav, fig.align= "center", message = FALSE, warning = FALSE}
qc_missed_cleavages(
  DIA_clean_uniprot,
  condrep,
  pep_grouping_key,
  pep_nr_of_missed_cleavages,
  fg_quantity,
  method = "intensity",
  plot = TRUE,
  interactive = FALSE
)

qc_missed_cleavages(
  DIA_clean_uniprot,
  condrep,
  pep_grouping_key,
  pep_nr_of_missed_cleavages,
  fg_quantity,
  method = "count",
  plot = TRUE,
  interactive = FALSE
)
```

### Peptide types

Peptide type (tryptic, semi-tryptic, non-tryptic) distributions are even throughout the different samples.

```{r quality_control_pep_type, fig.align = "center", message = FALSE, warning = FALSE}

qc_peptide_type(
  DIA_clean_uniprot,
  condrep,
  pep_grouping_key,
  pep_type,
  intensity = fg_quantity,
  method = "count",
  plot = TRUE,
  interactive = FALSE
)

qc_peptide_type(
  DIA_clean_uniprot,
  condrep,
  pep_grouping_key,
  pep_type,
  intensity = fg_quantity,
  method = "intensity",
  plot = TRUE,
  interactive = FALSE
)

```

### Number of peptide IDs per sample

The numbers of identified peptides are consistent throughout the different samples.

```{r quality_control_pep_ID, fig.align = "center", message = FALSE, warning = FALSE}

DIA_raw$condrep <- paste(DIA_raw$r_condition, DIA_raw$r_replicate, sep = "_")


qc_ids(DIA_raw, condrep, pep_grouping_key, condition = r_condition, intensity = fg_quantity)

```

### Principal component analysis (PCA)

```{r PCA, message = FALSE, warning = FALSE}

DIA_clean_uniprot %>%
qc_pca(condrep, 
       pep_grouping_key, 
       intensity_log2, 
       r_condition)

```

